Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers

### 2023-04-25 manuscript/manuscript.Rmd
updated rmarkdown manuscript according to the review

---
title             : "A survey of registration practices among observational researchers using preexisting datasets"
shorttitle        : "Explore and Confirm Analysis Workflow"
author:
  - name: Robert T. Thibault
    affiliation: '1,5'
    role:
      - Conceptualization
      - Data curation
      - Formal analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Supervision
      - Validation
      - Visualization
      - Writing - original draft
      - Writing - review & editing
    corresponding: yes
    email: robert.thibault@stanford.edu
    address: Enter postal address here
  - name: Marton Kovacs
    affiliation: '2,6'
    role:
      - Data curation
      - Formal analysis
      - Software
      - Validation
      - Visualization
      - Writing - review & editing
    email: marton.balazs.kovacs@gmail.com
  - name: Tom E. Hardwicke
    affiliation: '3'
    role:
      - Methodology
      - Writing - review & editing
  - name: Alexandra Sarafoglou
    affiliation: '4'
    role:
      - Methodology
      - Writing - review & editing
  - name: John P. A. Ioannidis
    affiliation: '4'
    role:
      - Methodology
      - Writing - review & editing
  - name: Marcus R. Munafò
    affiliation: '1,7'
    role:
      - Conceptualization
      - Methodology
      - Supervision
      - Writing - review & editing
affiliation:
  - id: '1'
    institution: Meta-Research Innovation Center at Stanford (METRICS), Stanford University.
  - id: '2'
    institution: Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest,
      Hungary
  - id: '3'
    institution: Melbourne School of Psychological Sciences, University of Melbourne.
  - id: '4'
    institution: Department of Psychology, University of Amsterdam.
  - id: '5'
    institution: School of Psychological Science, University of Bristol.
  - id: '6'
    institution: Institute of Psychology, ELTE Eotvos Lorand University, Budapest,
      Hungary
  - id: '7'
    institution: Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for
      Transforming Biomedical Research, Berlin Institute of Health, Charité – Universitätsmedizin
      Berlin.
  - id: '8'
    institution: MRC Integrative Epidemiology Unit at the University of Bristol.
  - id: '9'
    institution: Departments of Medicine, Epidemiology and Population Health, Biomedical
      Data Science, and Statistics, Stanford University.
abstract: |
  placeholder for an abstract

keywords          : "keywords"
wordcount         : "X"
# bibliography      : "references.bib"
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
appendix: "supplementary_materials.Rmd"

classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
# The developer version is needed
library(likert)
library(knitr)
library(viridis)
# r_refs("references.bib")
```

@@

# Introduction

# Methods

# Results
## Participants

@@

We invited the ALSPAC mailing list to participate, which included 1148 email addresses. 54 emails bounced, leaving 1094 emails that went through. The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.[^1] The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`).

Respondents published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$n_studies, .75, na.rm = TRUE), 1)`) studies using preexisting observational data (Figure S1). They reported using the programming languages `r language_text` (Table S1)[^2]. `r pull(filter(concerned, support == "positive"), percentage)`% (`r pull(filter(concerned, support == "positive"), n)`/`r pull(filter(concerned, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data (Figure C2); `r pull(filter(concerned, support == "negative"), percentage)`% (`r pull(filter(concerned, support == "negative"), n)`/`r pull(filter(concerned, support == "negative"), n_sum)`) reported being less concerned.

[^1]: The ALSPAC mailing list has been active for >30 years and may contain email addresses that are no longer monitored. For example, we received one email reply stating that the recipient hasn’t been active in research for 30 years. Excluding these email addresses would increase the response rate, but we do not know by how much.
[^2]: Participants could select multiple responses to this survey question.
[^3]: The survey defined trustworthy as: “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust.”
[^4]: The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.”

@@

Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy[^3] (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible[^4] (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 2, top panel). At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 2, bottom panel).

@@

(ref:typicallyEcawPlotCaption) __Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust”. The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.” For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. For the bottom panel we excluded the missing responses (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "Missing"), n)`) and responses of “I don't understand the question” (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "I don't understand the question"), n)`).

@@

Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Figure 2A). About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%; `r filter(script, method_script == "Never or almost never") %>% pull(n)`/`r filter(script, method_script == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%; `r filter(script, method_script == "Sometimes") %>% pull(n)`/`r filter(script, method_script == "Sometimes") %>% pull(n_sum)`) (Figure 3). `r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% (`r filter(blind, method_blind == "Never or almost never") %>% pull(n)`/`r filter(blind, method_blind == "Never or almost never") %>% pull(n_sum)`) reported that they never or almost never blind the data analyst (Figure 3). Almost all respondents answered that they use both exploratory (`r pull(exploratory, percentage)`%; `r pull(exploratory, n)`/`r pull(exploratory, n_sum)`) and confirmatory (`r pull(confirmatory, percentage)`%; `r pull(confirmatory, n)`/`r pull(confirmatory, n_sum)`) analyses at least sometimes (Figure 3).

@@

(ref:methodPlotCaption) __Responses to survey questions about the research practices of participants.__

@@

(ref:alspacPlotCaption) __Responses to survey questions about using ECAWs.__ For the 4 questions we excluded missing values (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Missing"), n)`), responses of "I don't understand the question" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of "Unsure" (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`).

@@

Table 1. Recurring topics in responses to the open-ended survey questions. The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments. These questions received a total of (`r nrow(open_ended)`) responses from (`r nrow(distinct(open_ended, response_id))`) unique respondents. A complete list of responses are viewable in the open data [LINK]. We synthesized the response to open-ended questions into the 9 topics on the left side of this table. We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as typical preregistration and ECAWs. On the right side of the table, we provide a reflection on each topic.

## Exploratory analyses

# Discussion

# Acknowledgements

>>>>

---
title             : "Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers"
shorttitle        : "Explore and Confirm Analysis Workflow"
author:
  - name: Robert T. Thibault
    affiliation: '1,2'
    corresponding: yes
    email: robert.thibault@stanford.edu
    address: Enter postal address here
  - name: Marton Kovacs
    affiliation: '3,4'
    email: marton.balazs.kovacs@gmail.com
  - name: Tom E. Hardwicke
    affiliation: '5'
  - name: Alexandra Sarafoglou
    affiliation: '6'
  - name: John P. A. Ioannidis
    affiliation: '1,7'
  - name: Marcus R. Munafò
    affiliation: '2,8'
affiliation:
  - id: '1'
    institution: Meta-Research Innovation Center at Stanford (METRICS), Stanford University.
  - id: '2'
    institution: School of Psychological Science, University of Bristol.
  - id: '3'
    institution: Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary
  - id: '4'
    institution: Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary
  - id: '5'
    institution: Melbourne School of Psychological Sciences, University of Melbourne.
  - id: '6'
    institution: Department of Psychology, University of Amsterdam.
  - id: '7'
    institution: Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for
      Transforming Biomedical Research, Berlin Institute of Health, Charité – Universitätsmedizin
      Berlin.
  - id: '8'
    institution: MRC Integrative Epidemiology Unit at the University of Bristol.
abstract: |
  \textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis. An alternative method that affords additional protection from researcher bias, which cannot be gained from conventional forms of preregistration alone, is an Explore and Confirm Analysis Workflow (ECAW). In this workflow, a data management organization initially provides access to only a subset of their dataset to researchers who request it. The researchers then prepare an analysis script based on the subset of data, upload the analysis script to a registry, and then receive access to the full dataset. ECAWs aim to achieve similar goals as preregistration, but make access to the full dataset contingent on compliance. The present survey aimed to garner information from the research community where ECAWs could be applied—employing the Avon Longitudinal Study of Parents and Children (ALSPAC) as a case example. \hfill \break \hfill \break \textbf{Methods.} We emailed a web-based survey to researchers who had previously applied for access to ALSPAC’s transgenerational observational dataset.\hfill \break \hfill \break \textbf{Results}. We received 107 responses, for a 10% response rate. The results suggest that—at least among our sample of respondents—ECAWs hold the potential to serve their intended purpose and appear relatively acceptable. For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs (versus 55% who agreed). However, as many as 25% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 44% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions. \hfill \break \hfill \break \textbf{Preregistration:} https://osf.io/g2fw5 Deviations from the preregistration are outlined in Supplementary Material A.

keywords          : "Blind data analysis; preregistration; ALSPAC; meta-research; open science; Explore and Confirm Analysis Workflow (ECAW)"
wordcount         : "3546"
bibliography      : "ecaw-ref.bib"
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
appendix: "supplementary_materials.Rmd"

classoption       : "man"
output            :
  papaja::apa6_pdf:
    latex_engine: xelatex
    extra_dependencies: "tcolorbox"
header-includes   :
  - \usepackage{tcolorbox}
  - \usepackage{caption}
  - \captionsetup[figure]{labelformat=empty}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
# The developer version is needed
library(likert)
library(knitr)
library(viridis)
r_refs("ecaw-ref.bib")
```

@@

# Introduction
Many published research findings are non-reproducible and potentially false or misleading [@camerer_evaluating_2016; @errington_investigating_2021; @ioannidis_why_2005; @ioannidis_why_2008; @collaboration_estimating_2015]. These shortcomings may stem from a combination of researcher bias, publication bias, selective reporting of results, incomplete reporting of methods, and other questionable research practices. These issues can lead to research waste, and ineffective or even harmful healthcare and policy interventions (e.g., see @prasad_ending_2019).

Some research disciplines have adopted standards to address these issues. For example, researchers conducting clinical trials regularly register outcome measures before enrolling participants, and they keep both participants and experimenters blind regarding group assignment.  In cases where datasets already exist, however, researchers cannot easily adopt these practices. Researchers performing secondary data analyses generally have access to complete datasets without any blinding imposed on them. This may lead to exploration of diverse analyses and selective reporting based on the nature of the results.

## Preregistration in the context of secondary data analyses
Increasing the uptake and effectiveness of preregistration for secondary data analyses may require a different approach than it does for clinical trials. For example, the analytical space for secondary data analyses (e.g., of a longitudinal cohort dataset) is more vast than the analytical space for most clinical trials. Thus, the clinical trial standard of registering outcome measures, but no analysis plan, remains insufficient. Moreover, preregistration of secondary data analyses remains uncommon, and unlike for clinical trials (where the dataset doesn’t yet exist), there is no guarantee that researchers performing secondary data analyses haven’t accessed the dataset. To extend the benefits of preregistration to secondary data analyses, we propose a workflow that necessitates the preregistration of an executable analysis script before researchers can access the full dataset.

## Explore and Confirm Analysis Workflow (ECAW)
In this intervention we propose, a data management organization that controls access to a preexisting data set would first provide access to only a subset of the data a researcher requests. After the researcher prepares an analysis script based on the subset of data and uploads it to a registry—where it will be openly accessible and permanent (e.g., to the Open Science Framework Registry)—the data management organization then provides access to the full dataset (e.g., via a secure data environment) and the researcher can proceed as they wish. The exact implementation (e.g., whether the data management organization runs quality checks on the analysis scripts) would depend on the preferences of the data management organization and the research community using their dataset (see Box 1 for an hypothetical example).

We call this process an Explore and Confirm Analysis Workflow (ECAW). Researchers may use the subset of data to generate hypotheses and/or to simply ensure that their intended analysis runs properly. ECAWs may increase the uptake of preregistration and provide assurance that the analyses were developed before the researchers observed the full dataset. Moreover, whereas conventional preregistrations often leave many degrees of freedom regarding exactly how the data will be analyzed (e.g., @bakker_ensuring_2020), ECAWs provide an executable analysis script. Researchers may also find ECAWs more agreeable than conventional preregistration because they allow for exploration. ECAWs would not fully solve publication bias, but could make it easier to detect because the registered analysis script would serve as a form of preregistration.

In short, ECAWs give researchers access to enough data to develop an informed analysis, but without compromising the confirmatory nature of a final analysis on the complete dataset.

\pagebreak

\begin{tcolorbox}
\textbf{Box 1. A hypothetical example of an ECAW in practice.}

\includegraphics[width=1\textwidth]{./ecaw_workflow.png}
{\fontsize{10pt}{10pt}\selectfont
The overarching framework for ECAWs is depicted above. Within this framework, data management organizations would need to pre-specify several details for each step. Below, we provide a hypothetical example for illustrative purposes. The exact implementation of ECAWs would depend on the preferences of each data management organization and the community who uses their dataset.

\begin{enumerate}
  \item A research team submits to a data management organization (i) a paragraph describing the analyses they want to run, and (ii) a list of the variables they want to analyze.
  \item The data management organization provides the research team with access to a subset of the data. For example, if the research team requests data for 15 variables collected from 10,000 participants, the data management organization will provide access to all 15 variables from a random subset of participants—say 1000.
  \item The research team prepares an analysis script written in a common programming language (e.g., R or Stata) using the subset of data. They register this analysis script, the output from the script, and the paragraph they sent to the data management organization in Step 1, to www.osf.io/registries as an ‘Open-Ended Registration’. The researchers are free to run as many analyses as they would like on the subset of data. However, they should only register the analysis script they plan to use on the complete dataset.
  \item The data management organization performs a basic quality check. They ensure that (i) the analysis script, the script output, and the paragraph are registered, and (ii) the output contains a clear number of itemized results (e.g., as clinicaltrials.gov does for outcome measures). If the quality check fails, the data management organization asks the researchers to update their registration until it passes. The research team is then given access to a dataset with all the variables requested for all participants.
  \item The researchers can proceed as they wish. They can run the registered analysis script on the complete dataset, make adjustments to the analysis if desired, or not proceed at all. The data management organization does not perform any additional check on what analyses were run. Regardless of what the research team decides, a permanent version of the planned analysis is available on the OSF Registry.
\end{enumerate}
}
\end{tcolorbox}

\pagebreak

## Examples of ECAW-like workflows
One study probed the benefits and drawbacks of a workflow similar to ECAWs, but they provided a synthetic dataset rather than a data subset (@sarafoglou_comparing_2023). The researchers recruited 120 teams to analyze a single observational dataset and had half the teams preregister a written analysis plan and the other half prepare an analysis plan by drafting an analysis script based on a dataset with shuffled data for the variables of interest (i.e., the analysts were effectively blinded). Based on self-reports from the participants, the researchers found that two workflows were comparable in terms of effort and that teams using blinded data analysis made fewer deviations from their analysis plan.

Some disciplines, such as particle physics, implement blind data analyses regularly (@klein_blind_2005; @maccoun_blind_2015). A version of the ECAW workflow has also been successfully implemented by eight teams performing secondary data analysis on a dataset managed by the Psychological Science Accelerator group (@Forscher2020) and is currently being used for Registered Reports based on a large psychology dataset (@Schmidt2023).

In medicine and clinical epidemiology, a similar workflow is implemented by the software platform OpenSAFELY (www.opensafely.org). This platform provides a dataset of simulated health records from which researchers can develop an analysis script. When ready, a researcher submits their analysis script which is automatically logged and made public in GitHub. The analysis then runs in a Trusted Research Environment (TRE) on data which is stored in the data centers where patients’ records already reside (i.e., the data is not copied or moved). This workflow keeps individual health records hidden while also documenting all analyses run on the real data.

## Study objectives
Here, we present a descriptive and exploratory survey study. We had no hypotheses, but we did have two specific objectives. (1) To gain insights on the opinions and practices of researchers who already use preexisting observational datasets, in regards to the trustworthiness and reproducibility of research. (2) To use these insights to inform future research—including a potential trial of ECAWs with the Avon Longitudinal Study of Parents and Children (ALSPAC)—on how data management organizations can encourage rigorous and reproducible research practices.

# Methods
## Participants
We sent an email to invite researchers on the mailing list for the UK-based Avon Longitudinal Study of Parents and Children (ALSPAC) to participate in an online survey (see Supplementary Material B). We partnered with ALSPAC because they manage an oft-requested dataset and expressed interest in studying ways to ensure the research stemming from their dataset is rigorous. ALSPAC is “a transgenerational prospective observational study investigating influences on health and development across the life course. It considers multiple genetic, epigenetic, biological, psychological, social and other environmental exposures in relation to a similarly diverse range of health, social and developmental outcomes.” (Boyd et al., 2013; Fraser et al., 2013). Thus, our invitation reached researchers that use observational data across the health and social sciences.  The survey was open from 10 Oct 2022 to 1 Nov 2022. We sent two reminder emails, exactly one week and two weeks after the original email invitation. The mailing list comprises researchers whose email addresses were present on a proposal to access the ALSPAC dataset and included 1148 email addresses.

## Survey
The survey contained 6 blocks and is available at [https://osf.io/5h7gb](https://osf.io/5h7gb). We developed the survey with feedback from the Principal Investigator and Executive Director of ALSPAC (Nicholas Timpson and Kate Northstone). We aimed to include as few questions as possible (to encourage a high response rate) while still garnering substantive information on whether ECAWs are relevant and acceptable to ALSPAC users.

Block 1 assessed the extent to which respondents believe that observational research using preexisting data is trustworthy and reproducible (2 questions). Block 2 asked respondents how often they use practices related to transparency and reducing researcher bias, including preregistration, blinded data analysis, and sharing analysis scripts (5 questions). Between Block 2 and Block 3, the survey described ECAWs. Block 3 assessed the extent to which respondents believe that ECAWs would make observational research using preexisting data more trustworthy and reproducible (2 questions). Block 4 directly asked respondents whether ALSPAC should run a study on ECAWs and whether they would participate (5 questions). Block 5 contained open-ended questions about the perceived benefits and drawbacks of ECAWs, reservations about ECAWs, whether additional incentives might be needed to use ECAW, and suggestions for other research practices or policies that data management organizations could implement to improve research quality (3 questions). Block 6 asked participants how concerned they are about research quality, how many relevant studies they have published, what software they use for data analysis, and for any additional comments (4 questions). All rating-scale questions contained the response option _“I don’t understand the question”_. The questions presented in Figure 3 also contained the response option _“Unsure”_.

## Analyses
We present the results for closed-ended questions as prevalence rates or counts in Figures 1-3 and Supplementary Material C. When reporting percentages, we collapse together all positive responses on the rating scales (e.g., _“strongly agree”_ and _“somewhat agree”_) as well as all negative responses. The total number of responses differ among questions due to the missing values and responses of _“I don’t understand the question”_ and _“Unsure”_. We narratively synthesize the responses to the open-ended questions in Table 1 (i.e., we provide a non-systematic qualitative summary).

# Results
## Survey completion
Of 1148 emails sent, 1094 went through and 54 bounced. The survey was completed `r nrow(processed)` times and partially completed `r nrow(dplyr::filter(processed, has_missing)) ` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(filter(processed, has_missing)) / 1094 * 100)`% for incomplete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to. The ALSPAC team only removes email addresses from their mailing list if someone explicitly requests this action. Thus, their list likely includes several email addresses that are no longer monitored. For example, we received one email reply stating that the recipient hasn’t been active in research for 30 years. It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions). These two factors may have deflated the response rate.} The median time taken for complete survey responses was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`). This manuscript presents the results for complete surveys. Supplementary Material D presents the results with partially complete surveys included.

## Participants

@@

Respondents had published a median of `r median(processed$n_studies)` (IQR `r round(quantile(processed$n_studies, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$n_studies, .75, na.rm = TRUE), 1)`) studies using preexisting observational data (Supplementary Figure C1). They reported using the following programming languages or software packages: `r language_text` (Supplementary Table C1)\footnote[2]{Participants could select multiple responses to this survey question.}. `r pull(filter(concerned, support == "positive"), percentage)`% (`r pull(filter(concerned, support == "positive"), n)`/`r pull(filter(concerned, support == "positive"), n_sum)`) of participants reported being more concerned with research trustworthiness, bias, rigour, and reproducibility compared to what they think of as a typical research who uses preexisting observational data; `r pull(filter(concerned, support == "negative"), percentage)`% (`r pull(filter(concerned, support == "negative"), n)`/`r pull(filter(concerned, support == "negative"), n_sum)`) reported being less concerned (Supplementary Figure C2).

@@

Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 2, top panel). At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).

@@

(ref:typicallyEcawPlotCaption) __Figure 1. Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust”. The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.” For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. The bar charts in the top panel had no missing responses or selection of the option _“I don’t understand the question”_. The bar charts in the bottom panel excluded missing responses (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "Missing"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "Missing"), n)`) and responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing, item == "Trustworthy" & value == "I don't understand the question"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible" & value == "I don't understand the question"), n)`).

@@

Over half of respondents reported that their studies using preexisting observational data are preregistered never or almost never (`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(percentage)`%; `r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n)`/`r filter(preregistered, method_preregistered == "Sometimes") %>% pull(n_sum)`) (Figure 2). About half reported sharing their analysis scripts never or almost never (`r filter(script, method_script == "Never or almost never") %>% pull(percentage)`%; `r filter(script, method_script == "Never or almost never") %>% pull(n)`/`r filter(script, method_script == "Never or almost never") %>% pull(n_sum)`), or sometimes (`r filter(script, method_script == "Sometimes") %>% pull(percentage)`%; `r filter(script, method_script == "Sometimes") %>% pull(n)`/`r filter(script, method_script == "Sometimes") %>% pull(n_sum)`). `r filter(blind, method_blind == "Never or almost never") %>% pull(percentage)`% (`r filter(blind, method_blind == "Never or almost never") %>% pull(n)`/`r filter(blind, method_blind == "Never or almost never") %>% pull(n_sum)`) reported that they never or almost never blind the data analyst. Almost all respondents answered that they use both exploratory (`r pull(exploratory, percentage)`%; `r pull(exploratory, n)`/`r pull(exploratory, n_sum)`) and confirmatory (`r pull(confirmatory, percentage)`%; `r pull(confirmatory, n)`/`r pull(confirmatory, n_sum)`) analyses at least sometimes.

@@

(ref:methodPlotCaption) __Figure 2. Responses to survey questions about the research practices of participants.__

@@

(ref:alspacPlotCaption) __Figure 3. Responses to survey questions about using ECAWs.__ These bar charts exclude missing values (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Missing"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Missing"), n)`; respectively from top to bottom), responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`). Agreement with the first question may be slightly inflated due to the format of the questions in this block. Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions. Four respondents agreed with all four statements, suggesting they may have glazed over the word “less” in the first question \protect\footnotemark[3]. Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study” \protect\footnotemark[4].

\protect\footnotetext[3]{Another four respondents agreed with at least the first and second question, which appear contradictory. We did not preregister these considerations. More careful wording of these questions could have circumvented the ambiguity in interpreting these seemingly contradictory responses.}
\protect\footnotetext[4]{We intended for these questions to ask about ALSPAC running a trial on ECAWs. However, due to the ambiguity around the terms “study”, some respondents may have interpreted this as a survey, focus group, feasibility, or pilot type of study.}

@@

__Table 1. Recurring topics in responses to the open-ended survey questions.__ The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments. These questions received a total of (`r nrow(open_ended)`) responses from (`r nrow(distinct(open_ended, response_id))`) unique respondents. A complete list of responses are viewable in the open data [LINK]. We synthesized the responses to open-ended questions into the 9 topics on the left side of this table. We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as conventional preregistration and ECAWs. On the right side of the table, we provide our reflection on each topic.

```{r openEndedTable, echo=FALSE, message=FALSE, warnings=FALSE, fig.align = "center", out.width="100%", fig.id=FALSE}
open_ended_table <- tibble(
  subheader = c(
    "Concerns about acceptability",
    "Concerns about acceptability",
    "Concerns about acceptability",
    "Concerns about acceptability",
    "Concerns about impact",
    "Concerns about impact",
    "Alternative interventions",
    "Alternative interventions",
    "Alternative interventions"
    ),
  `Respondents’ comments` = c(
      "1. ECAWs would take too much time.",
      "2. The subset of data would be insufficient for certain applications (e.g., for genome-wide association studies, rare outcomes, questions regarding small sub-samples, and imputation).",
      "3. Researchers will often have to change their analysis after seeing the full dataset.",
      "4. Researchers sometimes reuse the same dataset which they will already have full access to.",
      "5. Researchers can still p-hack / data-dredge when using ECAWs.",
      "6. Researchers may give findings from the subset too much weight (e.g., by not performing an analysis on the full dataset because it was null in the subset, or vice versa).",
      "7. Data management organizations should require final analysis scripts to be shared (so that results are reproducible, and easy for other researchers to build on).",
      "8. Data management organizations should maintain a repository that outlines each research project conducted using their dataset (to reduce duplication and facilitate replication).",
      "9. Data management organizations could make a synthetic version of their dataset openly available. That dataset could be used for ECAWs."
      ),
    `Our reflection` = c(
